---
title: "Visualizations"
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, cache = T)
```

```{r}
library(MixSim)
library(RColorBrewer)
library(MASS)
library(freqparcoord)
library(expm)
library(grDevices)
library(radviz3d)
library(rgl)

setwd("~/Desktop/temp/radviz3d-jcgs-codes/")
source("modradviz2d-2.R")

viz2d_3d <- function(x, angles =  2*pi*(0:(ncol(x)-1))/ncol(x)) {
  y <- apply(X = x, MARGIN = 2, FUN = function(z)((z - min(z))/(max(z) - min(z))))
  sy <- rowSums(y)/ncol(y)
  y <- apply(X = y, MARGIN = 1, FUN = function(z)(z/sum(z)))
  y <- t(y)
  pr <- cbind(cos(angles), sin(angles))
  xproj <- cbind(y %*% pr, sy)
}

# {viz3d} function based on Artero and Oliveira's paper.
viz3d <- function(x, angles =  2*pi*(0:(ncol(x)-1))/ncol(x)) {
  y <- apply(X = x, MARGIN = 2, FUN = function(z)((z - min(z))/(max(z) - min(z))))
  #    x <- apply(X = y, MARGIN = 1, FUN = function(z)(z/sum(z)))
  #    x <- t(x)
  pr <- cbind(cos(angles), sin(angles), 1)
  xproj <- y %*% pr / ncol(y)
}

# color palette
wine.palette = brewer.pal(8,"Dark2")[c(1,2,3)]
cb.palette = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00") 
oil.palette = c(brewer.pal(8,"Dark2"), "cyan")
# for SARS-COV2 data
kols1 <- c(RColorBrewer::brewer.pal(name = "Set2", n = 8)[-1],
           RColorBrewer::brewer.pal(name = "Greens", n = 9)[7],
           RColorBrewer::brewer.pal(name = "Reds", n = 9)[7],
           RColorBrewer::brewer.pal(name = "Blues", n = 9)[7])
kols2 <- c(RColorBrewer::brewer.pal(name = "Dark2", n = 8),
                  RColorBrewer::brewer.pal(name = "Reds", n = 9)[8],
                  RColorBrewer::brewer.pal(name = "Blues", n = 9)[8])
kols <- kols2

```

```{r}
##
## Display the map of pairwise overlap measures of Maitra and Melnykov
## (JCGS, 2012)
##
## overlap.mat = matrix of total pairwise overlaps (necessarily symmetric)
## map.col = colormap for the mapping
## linescol = color for the lines drawing the squares
## map.range = range of the overlap map (default: minimum and maximum of lower
##               triangle of the matrix)
## lab.col = color of the labels (same as nrow(matrix) if provided)
## lab.cex = character size of the label
## map.cex = character size of the overlap values laid on the map
## legend.cex = character size of the legend text (does not work always)
##
## provides map of overlap values for each group of mixture model
##
## written Ranjan Maitra, Ames, IA 50011-1210, June 28, 2009
##
## modified Ranjan Maitra, Ames, IA 50011-1090, December 23, 2016.
## last modified Ranjan Maitra, Ames, IA 50011-1090, October 31, 2020.
##
## modification to bring in specifications for label color, labels, maps and
## legend character size; subsequently change default color of map and lines
##

overlap.map <-
    function(overlap.mat, map.col = RColorBrewer::brewer.pal(name = "PuRd", n = 9), linescol = "#1D91C0", map.range = NULL, lab = NULL, lab.col = 1, lab.cex = 5, font = 1, map.cex = 5, legend.cex = 1, scale = NULL, scale.pos = -2, legend.width = 1)
{
    oxmat <- overlap.mat
    oxmat[lower.tri(oxmat)] <- NA
    diag(oxmat) <- NA
    p <- ncol(oxmat)
    if(is.null(lab)){
      lab <- 1:p
    }
    newox <- oxmat[-p, -1]
    newox <- cbind(rbind(NA, newox), NA)[, p:1]

    layout(matrix(c(rep(1, 4*p^2), rep(2, 2*p), rep(3, 2*p), rep(4, 2*p), rep(4, 2*p)), nrow = 2*p, ncol = 2*p + 4 ))
    ##  layout(matrix(c(rep(1, p^2), rep(2, p), rep(3, p), rep(4, p)), nrow = p, ncol = p + 3))

    
    par(mar = c(0.1,0.1,0.75,0.1))
    if (is.null(map.range)) map.range <- range(newox, na.rm = T)  
    image(x = 1:p, y = 1:p, z = newox, axes = F, xlab = "", ylab = "", #col = brewer.pal(9, "GnBu"))
          col = map.col, zlim = map.range)
    text(y = 2:p, x = rep(1, p-1), labels = lab[p:2], cex = lab.cex, font = font, col = lab.col[p:2])
    text(x = 2:p, y = rep(1, (p-1)), labels = lab[1:(p-1)], cex = lab.cex, font = font, col = lab.col[1:(p-1)])
    if (!is.null(scale)){
       text(x = p-scale.pos, y = p+0.2, labels = scale, cex = map.cex)
    }
   


    for(i in 1:p) {
        for(j in i:p) {
            lines(x = c(i+0.5, i+0.5), y = c(p-j+1,p-j)+1.5, col = linescol, lwd = 0.5)
            lines(y = c(i+0.5, i+0.5), x = c(p-j+1,p-j)+1.5, col = linescol, lwd = 0.5)
        }
    }
    for(i in 2:p) text(x=1:p, y = i, labels = round(newox[,i], 2),
                       col = ifelse(newox[,i] < median(map.range), "black", "white"), cex = map.cex)

   frame()

    # savepar <- par(cex=0.75, lwd=0.25, mar = c(1, 0.5, 1, 2),
    #                 xaxs="i", yaxs="i")
    
    savepar <- par(cex=0.75, lwd=0.25, mar = c(0.5, 0.5, 0.5, 0.5),
                   xaxs="i", yaxs="i")
    # if (legend){
    plot.new()
    length.col <- length(map.col) + 1
    ra <- seq(from = map.range[1], to = map.range[2], length=length.col)
    plot.window(xlim=c(0,0.1), ylim= c(map.range[1], map.range[2]))
    rect(0, ra[-length.col], legend.width, ra[-1], col = map.col, border = NULL)
    axis(4, at = ra, labels = round(ra, digits = 2), las = 1, cex.axis = legend.cex, line = NA)
    

    rect(0, 1, legend.width, ra[length.col], col = NULL)
    frame()
    }

```

We display visualizations using RadViz2D, RadViz3D, Viz3D for simulation datasets and real datasets.

# Simulated Datasets
```{r, eval = F}
omegas <- c(0.001, 0.01, 0.05)
sim_data <- NULL
overlap_mat_sim <- NULL

seed = c(456, 1234, 3456)
for (i in 1:length(omegas)){
  set.seed(seed[i])
  p = 5
  Q <- MixGOM(goMega = omegas[i], K = 5, p = p, sph = F, hom = F)
  overlap_mat_sim[[i]] <- overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap
  print(t(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap) + overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap - diag(5))
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  # colnames(sn) <- paste(1:4)
  class <- as.factor(A$id)
  sn.df <- data.frame(class,sn)
  sim_data[[i]] <- sn.df
}

save(overlap_mat_sim, sim_data, file = "sim_data_5.rda")
```


## Fig. S1: Overlap Map, RadViz2D, Viz3D and RadViz3D of Simulated 5D Datasets

(a) [RadViz2D](radviz2d_jcgs.html)
(b) [Viz3D](viz3d_jcgs.html)
(c) [RadViz3D](radviz3d_jcgs.html)


<!-- # ```{r} -->
<!-- # res_radviz <- list() -->
<!-- # set.seed(2015) -->
<!-- # for (i in 1:length(omegas)){ -->
<!-- #   sn.df <- sim_data[[i]] -->
<!-- #   radialvis3d(data = sn.df[,-ncol(sn.df)], cl = as.factor(sn.df[,ncol(sn.df)]), domrp = T, doGtrans = T, lwd = 2, alpha = 0.325, radius = 0.02, color = brewer.pal(5, "Paired"), message = F) -->
<!-- #   # rgl::par3d(windowRect = c(0, 0, 1200, 1200)) -->
<!-- #   # rgl::par3d(viewport = c(0,0,1200, 1200)) -->
<!-- #   # legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01)) -->
<!-- #   rgl::rgl.viewpoint(zoom = 0.6) -->
<!-- #   res_radviz[[i]] <- rgl::rglwidget();rgl.close() -->
<!-- # } -->
<!-- # ``` -->















# Real Datasets

## Fig. S2: Wine Dataset
```{r}
wine = read.table(file = "wine.dat", sep=",",
                  header=F, col.names = c("cultivar","Ahl","Malic","Ash", "Alk","Mg",
                                          "Phnls","Flvds","Nonfp","Pthyns",
                                          "Color","Hue","ODdil","Prol")) 
df = wine
```


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
df$cultivar = factor(df$cultivar)
N <- nrow(df)
p <- ncol(df) - 1
K <- nlevels(df$cultivar)
Pi <- rep(0, K)
Mu <- array(dim = c(K, p))
S <- array(dim = c(p, p, K))
level <- levels(df$cultivar)
for (i in 1:nlevels(df$cultivar)){
  Pi[i] <- sum(df$cultivar== level[i])/N
  S[,,i] <- cov(df[,-1][df$cultivar == level[i],])
  Mu[i,] <- colMeans(df[,-1][df$cultivar == level[i],])
}


overlap_res<- overlap(Pi = Pi, Mu = Mu, S = S)

overlap_map_sym <- t(overlap_res$OmegaMap) + overlap_res$OmegaMap - diag(K)

overlap.map(overlap_map_sym*1000, lab.col = wine.palette[1:K], scale = expression(""%*%10^-3),  map.cex = 2.75, lab.cex = 4, legend.cex = 1.5, font = 2,
            scale.pos = 0, legend.width = 0.5)
```
<br />
Overlap Map
</td>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-1], cl = df$cultivar, modify = F, 
            coord.labels = colnames(df[,-1]), class.labels = NULL,
            palette = wine.palette, opt.axis.order = F)
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-1])
res1 = data.frame(res1, as.factor(df$cultivar))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(xlab = "", ylab="", zlab="",axes = F, x,y,z, type="s",size = 0.95,
                  col = wine.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget();rgl.close()

```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-1], cl = factor(df$cultivar), domrp = F, doGtrans = F, lwd = 2, alpha = 0.02, radius = 0.035, 
            color = wine.palette, coord.font = 4,coord.cex = 1.5,
            class.labels = NULL)
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget();
rgl.close()
```
<br />
RadViz3D
</td>
</tr>
</table>

## Fig. S3: Crabs Dataset with CL
```{r}
data("crabs")
crabs$group = as.factor(crabs$sp:crabs$sex)
df = crabs[,4:9]
```


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
df$group = factor(df$group)
N <- nrow(df)
p <- ncol(df) - 1
K <- nlevels(df$group)
Pi <- rep(0, K)
Mu <- array(dim = c(K, p))
S <- array(dim = c(p, p, K))
level <- levels(df$group)
for (i in 1:nlevels(df$group)){
  Pi[i] <- sum(df$group== level[i])/N
  S[,,i] <- cov(df[,-6][df$group == level[i],])
  Mu[i,] <- colMeans(df[,-6][df$group == level[i],])
}


overlap_res<- overlap(Pi = Pi, Mu = Mu, S = S)

overlap_map_sym <- t(overlap_res$OmegaMap) + overlap_res$OmegaMap - diag(K)

overlap.map(overlap_map_sym*100, lab.col = cb.palette[1:K], scale = expression(""%*%10^-2~""), map.cex = 2.75, lab.cex = 4, legend.cex = 1.5, 
            font = 2, scale.pos = 0.1, legend.width = 0.1)
```
<br />
Overlap Map
</td>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-6], cl = df$group, modify = F, 
            coord.labels = colnames(df[,-6]), class.labels = NULL,
            palette = cb.palette, opt.axis.order = F)
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-6])
res1 = data.frame(res1, as.factor(df$group))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(xlab = "", ylab="", zlab="",axes = F, x,y,z, type="s",size = 0.75,
                  col = cb.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget();rgl.close()

```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-6], cl = factor(df$group), domrp = F, doGtrans = F, lwd = 2, alpha = 0.325, radius = 0.035, color = cb.palette, class.labels = NULL)
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget();rgl.close()
```
<br />
RadViz3D
</td>
</tr>
</table>

## Fig. S4: Crabs Dataset without CL
```{r}
data("crabs")
crabs$group = as.factor(crabs$sp:crabs$sex)
df = crabs[,4:9]
df = df[,-3]
```


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
df$group = factor(df$group)
N <- nrow(df)
p <- ncol(df) - 1
K <- nlevels(df$group)
Pi <- rep(0, K)
Mu <- array(dim = c(K, p))
S <- array(dim = c(p, p, K))
level <- levels(df$group)
for (i in 1:nlevels(df$group)){
  Pi[i] <- sum(df$group== level[i])/N
  S[,,i] <- cov(df[,-5][df$group == level[i],])
  Mu[i,] <- colMeans(df[,-5][df$group == level[i],])
}


overlap_res<- overlap(Pi = Pi, Mu = Mu, S = S)

overlap_map_sym <- t(overlap_res$OmegaMap) + overlap_res$OmegaMap - diag(K)

overlap.map(overlap_map_sym*100, lab.col = cb.palette[1:K], scale = expression(""%*%10^-2~""), map.cex = 2.75, lab.cex = 4, legend.cex = 1.5, 
            font = 2, scale.pos = 0.1, legend.width = 0.1)
```
<br />
Overlap Map
</td>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-5], cl = df$group, modify = F, 
            coord.labels = colnames(df[,-5]), class.labels = NULL,
            palette = cb.palette, opt.axis.order = F)
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-5])
res1 = data.frame(res1, as.factor(df$group))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(xlab = "", ylab="", zlab="",axes = F, x,y,z, type="s",size = 0.75,
                  col = cb.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget();rgl.close()

```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-5], cl = factor(df$group), domrp = F, doGtrans = F, lwd = 2, alpha = 0.325, radius = 0.035, color = cb.palette, class.labels = NULL)
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget();rgl.close()
```
<br />
RadViz3D
</td>
</tr>
</table>


## Fig. S5: Olive Oils Dataset with 9 Regions of Italy.
```{r}
data("oliveoils")
oliveoils = oliveoils[,-2]
class = factor(oliveoils$Area)
class = sub("North-Apulia","1-North-Apulia",class)
class = sub("South-Apulia","2-South-Apulia",class)
class = sub("Calabria","3-Calabria",class)
class = sub("Sicily","4-Sicily",class)
class = sub("Umbria","5-Umbria",class)
class = sub("East-Liguria","6-East-Liguria",class)
class = sub("West-Liguria","7-West-Liguria",class)
class = sub("Coast-Sardinia","8-Coast-Sardinia",class)
class = sub("Inland-Sardinia","9-Inland-Sardinia",class)
df = data.frame(class, oliveoils[,-1])

```


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
df$class = factor(df$class)
N <- nrow(df)
p <- ncol(df) - 1
K <- nlevels(df$class)
Pi <- rep(0, K)
Mu <- array(dim = c(K, p))
S <- array(dim = c(p, p, K))
level <- levels(df$class)
order <- c(8,9,5:7,1:4)
for (i in 1:nlevels(df$class)){
  j <- order[i]
  Pi[i] <- sum(df$class == level[j])/N
  S[,,i] <- cov(df[,-1][df$class == level[j],])
  Mu[i,] <- colMeans(df[,-1][df$class == level[j],])
}


overlap_res<- overlap(Pi = Pi, Mu = Mu, S = S)

overlap_map_sym <- t(overlap_res$OmegaMap) + overlap_res$OmegaMap - diag(K)

overlap.map(overlap_map_sym*1000, lab.col = oil.palette[order], 
            scale = expression(""%*%10^-3~""), map.cex = 2.55, lab.cex = 4, 
            font = 2, legend.cex = 2.1, scale.pos = 1.5, legend.width = 0.05)
```
<br />
Overlap Map
</td>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-1], cl = df$class, modify = F, 
            coord.labels = colnames(df[,-1]), class.labels = NULL,
            palette = oil.palette, opt.axis.order = F)
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-1])
res1 = data.frame(res1, as.factor(df$class))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(xlab = "", ylab="", zlab="",axes = F, x,y,z, type="s",size = 0.75,
                  col = oil.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget();rgl.close()

```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-1], cl = factor(df$class), domrp = F, doGtrans = F, lwd = 2, alpha = 0.325, radius = 0.035, color = oil.palette, coord.cex = 1.15,
            class.labels = NULL)
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget();rgl.close()
```
<br />
RadViz3D
</td>
</tr>
</table>

## Fig. S6: Olive Oils Dataset with 3 Areas of Italy.
```{r}
data("oliveoils")
oliveoils = oliveoils[,-2]
class = factor(oliveoils$Area)
class = sub("North-Apulia","1-North-Apulia",class)
class = sub("South-Apulia","2-South-Apulia",class)
class = sub("Calabria","3-Calabria",class)
class = sub("Sicily","4-Sicily",class)
class = sub("Umbria","5-Umbria",class)
class = sub("East-Liguria","6-East-Liguria",class)
class = sub("West-Liguria","7-West-Liguria",class)
class = sub("Coast-Sardinia","8-Coast-Sardinia",class)
class = sub("Inland-Sardinia","9-Inland-Sardinia",class)

group = class
group = sub("1-North-Apulia","1-south-italy",group)
group = sub("2-South-Apulia","1-south-italy",group)
group = sub("3-Calabria","1-south-italy",group)
group = sub("4-Sicily","1-south-italy",group)
group = sub("5-Umbria","2-north-italy",group)
group = sub("6-East-Liguria","2-north-italy",group)
group = sub("7-West-Liguria","2-north-italy",group)
group = sub("8-Coast-Sardinia","3-sardinia",group)
group = sub("9-Inland-Sardinia","3-sardinia",group)
df = data.frame(class = group, oliveoils[,-1])

```


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
df$class = factor(df$class)
N <- nrow(df)
p <- ncol(df) - 1
K <- nlevels(df$class)
Pi <- rep(0, K)
Mu <- array(dim = c(K, p))
S <- array(dim = c(p, p, K))
level <- levels(df$class)
for (i in 1:nlevels(df$class)){
  Pi[i] <- sum(df$class == level[i])/N
  S[,,i] <- cov(df[,-1][df$class == level[i],])
  Mu[i,] <- colMeans(df[,-1][df$class == level[i],])
}


overlap_res<- overlap(Pi = Pi, Mu = Mu, S = S)

overlap_map_sym <- t(overlap_res$OmegaMap) + overlap_res$OmegaMap - diag(K)

overlap.map(overlap_map_sym*1000, lab.col = wine.palette[1:K], 
            scale = expression(""%*%10^-3~""), map.cex = 2.75, lab.cex = 4, 
            font = 2, legend.cex = 1.5, scale.pos = 0, legend.width = 1)
```
<br />
Overlap Map
</td>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-1], cl = df$class, modify = F, 
            coord.labels = colnames(df[,-1]), class.labels = NULL,
            palette = wine.palette, opt.axis.order = F)
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-1])
res1 = data.frame(res1, as.factor(df$class))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(xlab = "", ylab="", zlab="",axes = F, x,y,z, type="s",size = 0.75,
                  col = wine.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget();rgl.close()

```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-1], cl = factor(df$class), domrp = F, doGtrans = F, lwd = 2, alpha = 0.325, radius = 0.035, color = wine.palette, coord.cex = 1.5,
            class.labels = NULL)
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget();rgl.close()
```
<br />
RadViz3D
</td>
</tr>
</table>

## Fig. S7: SARS-COV2 Dataset
```{r}
setwd("~/Desktop/temp/radviz3d-jcgs-codes/datasets/")
load("SARS-CoV-2-timeG12MA.rda")
dat.w.g12.ma$group <- factor(dat.w.g12.ma$group, 
                             levels =c("6/19/21", "6/26/21", "7/3/21", "7/10/21", 
                                       "7/17/21", "7/24/21", "7/31/21", "8/7/21", 
                                       "8/14/21", "8/21/21"))
dat.w.g12.ma$region = factor(dat.w.g12.ma$region)
df = dat.w.g12.ma
## Use hclust to find the order of coordinates for RadViz displays
#xx  <-  t(scale(asin(sqrt(dat.w.g12.ma[,-c(1:3,15)]))))
#x.hcl  <- hclust(dist(xx), method = "ward.D2")
#plot(x.hcl)
ll <- c()
ll[5] <- 5
ll[6] <- 4
ll[1:4] <- c(1, 7, 2, 8)
ll[7] <- 9
ll[8:11] <- c(3,10,6,11)
```

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
modradviz2d(x = df[,-c(1:3,15)][ll], cl = factor(df$region), modify = F, ret.proj = T,
            coord.labels = colnames(df[,-c(1:3,15)][ll]), class.labels = NULL, 
            palette = kols, pch = letters[1:10][df$group], opt.axis.order = F) -> xproj
new.dat <- data.frame(xproj, group = df$group, region = df$region)
for (i in 1:10) {
  dat <- new.dat[new.dat$region==i,]
  for (j in 1:11)
    lines(x = dat[j:(j+1), 1], y = dat[j:(j+1), 2], col = kols[i], lwd = 1.5)
}
```
<br />
RadViz2D
</td>
</tr>

<tr>

<td align="center" valign="center">
```{r}
res1 = viz2d_3d(df[,-c(1:3,15)][ll])
res1 = data.frame(res1, as.factor(df$region))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
pch3d(x = res1, y = NULL, z = NULL, xlab = "", ylab="", zlab="",axes = F,
      pch = letters[1:10][df$group],size = 0.95, color = kols)
new.dat <- data.frame(res1[,-4], group = df$group, region = df$region)
for (i in 1:10) {
  dat <- new.dat[new.dat$region==i,]
  for (j in 1:11)
    rgl.lines(x = dat[j:(j+1), 1], y = dat[j:(j+1), 2], z = dat[j:(j+1), 3], 
              color = kols[i], add = T, lwd = 1.5)
}
box3d()
rglwidget()
rgl.close()
```
<br />
Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = df[,-c(1:3)][ll], cl = factor(df$region), domrp = F, doGtrans = F, 
            lwd = 3, alpha = 0.025, pradius = 0.01, color = kols,
            with.class.labels = FALSE, shininess = 128, specular = 'black', back = "cull",
            ret.trans = TRUE, pch =  letters[1:12][dat.w.g12.ma$group.ma], point.cex = 1)-> dat.trns
new.dat <- data.frame(dat.trns, group = df$group, region = df$region)
for (i in 1:10) {
    dat <- new.dat[new.dat$region==i,]
    rgl.texts(x = dat[1, 1] + .05, y = dat[1, 2] + .05, z = dat[1, 3] + .05, 
                   text = as.character(i), cex = 3, color = kols[i])
    for (j in 1:11)
        rgl.lines(x = dat[j:(j+1), 1], y = dat[j:(j+1), 2], z = dat[j:(j+1), 3], 
                       color = kols[i], add = T, lwd = 2)
}
rgl.viewpoint(zoom = 0.6)
rglwidget()
```
<br />
RadViz3D
</td>
</tr>
</table>









