---
title: "Visualizations"
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, cache = T)
```

```{r}
source("stars.R")

viz2d_3d <- function(x, angles =  2*pi*(0:(ncol(x)-1))/ncol(x)) {
  y <- apply(X = x, MARGIN = 2, FUN = function(z)((z - min(z))/(max(z) - min(z))))
  sy <- rowSums(y)/ncol(y)
  y <- apply(X = y, MARGIN = 1, FUN = function(z)(z/sum(z)))
  y <- t(y)
  pr <- cbind(cos(angles), sin(angles))
  xproj <- cbind(y %*% pr, sy)
}

# {viz3d} function based on Artero and Oliveira's paper.
viz3d <- function(x, angles =  2*pi*(0:(ncol(x)-1))/ncol(x)) {
  y <- apply(X = x, MARGIN = 2, FUN = function(z)((z - min(z))/(max(z) - min(z))))
  #    x <- apply(X = y, MARGIN = 1, FUN = function(z)(z/sum(z)))
  #    x <- t(x)
  pr <- cbind(cos(angles), sin(angles), 1)
  xproj <- y %*% pr / ncol(y)
}

# color palette
load("colpal.rda")
dk.palette = senator.palette = c(rgb(red = 173, green = 216, blue = 230, max = 255), rgb(max = 255, red = 139, green = 0, blue = 0))

library(MixSim)
library(MASS)
library(expm)
library(grDevices)
library(freqparcoord)
library(Rtsne)
library(Rdimtools)
library(radviz3d)
library(rgl)
library(uwot)
library(RColorBrewer)

```

We display visualizations using Radviz3d, Viz3D, 3D UMAP, 3D tSNE and Star coordinate plots with OLDA and ULDA for simulated datasets and real datasets.

# Simulated Datasets

## Overlap of simulated groups
```{r echo=F}
library(MixSim)
omegas <- c(0.0001, 0.001, 0.01, 0.05, 0.25)
sim_data <- NULL

set.seed(2020)

for (i in 1:length(omegas)){
  p = 100
  Q <- MixGOM(goMega = omegas[i], K = 5, p = p, sph = T, hom = T)
  print(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap)
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  # colnames(sn) <- paste(1:4)
  class <- as.factor(A$id)
  sn.df <- data.frame(sn,class)
  sim_data[[i]] <- sn.df
}

convert_to_discrete <- function(data, discrete_p = 1:50){
  n <- nrow(data)
  data[,discrete_p] <- apply(data[,discrete_p], MARGIN = 2, FUN = function(x) floor(order(x)/(n+1)*10))
  return(data)
}


for (i in 1:5){
  sim_data[[i]] <- convert_to_discrete(sim_data[[i]])
}

```

```{r, eval = F}
save(sim_data, file = "datasets/sim_data.rda")
```

## Fig. S1: UMAP and RadViz3D of 100D MixSim Simulated Datasets

(a) [t-SNE](tsne.html)
(b) [UMAP](umap.html)
(c) [Stars with OLDA](stars_olda.html)
(d) [Stars with ULDA](stars_ulda.html)
(e) [Viz3D](viz3d.html)
(f) [RadViz3D](radviz3d.html)


<!-- # ```{r} -->
<!-- # set.seed(2020) -->
<!-- # res_umap_supervised <- list() -->
<!-- # for (i in 1:length(omegas)){ -->
<!-- #   sn.df <- sim_data[[i]] -->
<!-- #   class <- sn.df[,ncol(sn.df)] -->
<!-- #   umap_res <- umap(as.matrix(sn.df[,-ncol(sn.df)]), n_components = 3, y = sn.df[,ncol(sn.df)], target_weight = 0.5) -->
<!-- #   # umap_res <- umap(as.matrix(sn.g.df[,-20243]), n_components = 3) -->
<!-- #   res1 = data.frame(umap_res, class) -->
<!-- #   colnames(res1) = c("x","y","z","class") -->
<!-- #   # open 3d window -->
<!-- #    -->
<!-- #   # resize window -->
<!-- #   par3d(windowRect = c(100, 100, 612, 612)) -->
<!-- #   # plot -->
<!-- #   with(res1, plot3d(x,y,z, type="s",size = 0.75,  -->
<!-- #                     col = brewer.pal(5,"Paired")[as.numeric(class)])) -->
<!-- #   res_umap_supervised[[i]] <- rglwidget() -->
<!-- # } -->
<!-- #  -->
<!-- #  -->
<!-- # res_radviz <- list() -->
<!-- # set.seed(2015) -->
<!-- # for (i in 1:length(omegas)){ -->
<!-- #   sn.df <- sim_data[[i]] -->
<!-- #   radialvis3d(data = sn.df[,-ncol(sn.df)], cl = as.factor(sn.df[,ncol(sn.df)]), domrp = T, doGtrans = T, lwd = 2, alpha = 0.35, pradius = 0.02, color = brewer.pal(5, "Paired"), message = F) -->
<!-- #   # rgl::par3d(windowRect = c(0, 0, 1200, 1200)) -->
<!-- #   # rgl::par3d(viewport = c(0,0,1200, 1200)) -->
<!-- #   # legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01)) -->
<!-- #   rgl::rgl.viewpoint(zoom = 0.6) -->
<!-- #   res_radviz[[i]] <- rgl::rglwidget() -->
<!-- # } -->
<!-- # ``` -->















# Real Datasets

## Fig. S2: Gamma Ray Burst Dataset


<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/GRB_normal.Rda")
sn <- as.matrix(simd_normal_class[,-1])
class <- as.factor(simd_normal_class[,1])
sn.df <- data.frame(sn,class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.df[,-10]), dims = 3, 
              perplexity=min(500,floor((nrow(sn.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
umap_res <- umap(as.matrix(sn.df[,-10]), n_components = 3, y = sn.df[,10])
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()

```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
# stars with olda
out2 = do.olda(as.matrix(sn.df[,-10]), ndim = 4, class)
grb.olda = data.frame(out2$Y,class)
stars(grb.olda[,-5], grb.olda$class, col.palette = df.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
# stars with olda
out2 = do.ulda(as.matrix(sn.df[,-10]), ndim = 4, class)
grb.ulda = data.frame(out2$Y,class)
stars(grb.ulda[,-5], grb.ulda$class, col.palette = df.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
load("datasets/grbdf.rda")

res1 = viz2d_3d(sndf[,-ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[,ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>

<td align="center" valign="center">
```{r}
radialvis3d(data = sndf[,-ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = df.palette)
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>

## Fig. S3: Zipcode Digits Dataset

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/zip_normal.Rda") 
zip <- as.matrix(zip_normal_class[,-1])
class <- as.factor(zip_normal_class$class)
zip.df <- data.frame(zip,class)
set.seed(1)
tsne <- Rtsne(as.matrix(zip.df[,-257]), dims = 3, 
              perplexity=min(500,floor((nrow(zip.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = zip.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
set.seed(1)
umap_res <- umap(as.matrix(zip.df[,-257]), n_components = 3, y = zip.df[,257], target_weight = 0)
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = zip.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
<td>
</tr>


<tr>
<td align="center" valign="center">

```{r}
# stars with olda
out2 = do.olda(as.matrix(zip.df[,-257]), ndim = 4, class)
zip.olda = data.frame(out2$Y,class)
stars(zip.olda[,-5], zip.olda$class, col.palette = zip.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">

```{r}
# stars with ulda
out2 = do.ulda(as.matrix(zip.df[,-257]), ndim = 4, class)
zip.ulda = data.frame(out2$Y,class)
stars(zip.ulda[,-5], zip.ulda$class, col.palette = zip.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">

```{r}
load("datasets/zipdf.rda")
res1 = viz2d_3d(zipdigit[, -ncol(zipdigit)])
res1 = data.frame(res1, as.factor(zipdigit[, ncol(zipdigit)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = zip.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>

<td align="center" valign="center">
```{r}
load("datasets/zipdf.rda")
radialvis3d(data = zipdigit[, -ncol(zipdigit)], cl = zipdigit[,ncol(zipdigit)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = zip.palette)
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>

## Fig. S4: Faces

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/face_normal.Rda")
#get the cooridinate mean
y <- face_normal_class[,-1]
class <- as.factor(face_normal_class$class)
face.df <- data.frame(y,class)
g = c(1, 6, 8, 30, 33, 39)
face.g.df = face.df[face.df$class %in% g, ]
class = droplevels(face.g.df$class)
set.seed(1)
tsne <- Rtsne(as.matrix(face.g.df[,-281]), dims = 3, 
              perplexity=min(500,floor((nrow(face.g.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">

```{r}
umap_res <- umap(as.matrix(face.g.df[,-281]), n_components =  3, y = face.g.df[,281])

res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
out2 = do.olda(as.matrix(face.g.df[,-281]), ndim = 4, class)
face.olda = data.frame(out2$Y,class)
stars(face.olda[,-5], face.olda$class, col.palette = df.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
out2 = do.ulda(as.matrix(face.g.df[,-281]), ndim = 4, class)
face.ulda = data.frame(out2$Y,class)
stars(face.ulda[,-5], face.ulda$class, col.palette = df.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">

```{r}
load("datasets/facedf.rda")

res1 = viz2d_3d(facedf[, -ncol(facedf)])
res1 = data.frame(res1, as.factor(facedf[, ncol(facedf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>

<td align="center" valign="center">

```{r}
load("datasets/facedf.rda")
radialvis3d(data = facedf[, -ncol(facedf)], cl = facedf[,ncol(facedf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = df.palette, class.labels = c("A", "B", "C", "D", "E", "F"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
</br>
(f) RadViz3D
</td>
</tr>
</table>

## Fig. S5: Suicide Ideation
<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load('datasets-gdt/suicide_preprocess.rda')
sn.df <- suicide_id_normal
class <- as.factor(class)
sn.df <- data.frame(sn.df,class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.df[,-ncol(sn.df)]), dims = 3, 
              perplexity=min(500,floor((nrow(sn.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">

```{r}
set.seed(1)
umap_res <- umap(as.matrix(sn.df[,-ncol(sn.df)]), n_components =  3, y = sn.df[,ncol(sn.df)])
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<!-- <tr> -->
<!-- <td align="center" valign="center"> -->
<!-- ```{r} -->
<!-- out2 = do.olda(as.matrix(sn.df[,-ncol(sn.df)]), ndim = 4, class) -->
<!-- suicide.olda = data.frame(out2$Y,class) -->
<!-- stars(suicide.olda[,-5], suicide.olda$class, col.palette = df.palette) -->
<!-- ``` -->
<!-- <br /> -->
<!-- (c) Stars with OLDA -->
<!-- </td> -->

<!-- <td align="center" valign="center"> -->
<!-- ```{r} -->
<!-- out2 = do.ulda(as.matrix(sn.df[,-ncol(sn.df)]), ndim = 4, class) -->
<!-- suicide.ulda = data.frame(out2$Y,class) -->
<!-- stars(suicide.ulda[,-5], suicide.ulda$class, col.palette = df.palette) -->
<!-- ``` -->
<!-- <br /> -->
<!-- (d) Stars with ULDA -->
<!-- </td> -->
<!-- </tr> -->

<tr>
<td align="center" valign="center">
```{r}
sndf <- data.frame(suicide_prep, class)
res1 = viz2d_3d(sndf[, -ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[, ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = df.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(c) Viz3D
</td>


<td align="center" valign="center">
```{r}
# radialvis3d(data = sndf[, -ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.01, color = df.palette, class.labels = c("Non-suicical controls", "Suicidal attempters", "Suicical non-attempters"))

radialvis3d(data = sndf[, -ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = df.palette)
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(d) RadViz3D
</td>
</tr>
</table>

## Fig. S6: Voting Records of US Senators (Senators Dataset)

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/senators_normal.rda")
senators_normal_class[35,1] = "D"
senators_normal_class$party <- droplevels(senators_normal_class$party)
sn <- as.matrix(senators_normal_class[,-1])
class <- senators_normal_class$party
sn.df <- data.frame(sn,class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.df[,-543]), dims = 3, 
              perplexity=min(500,floor((nrow(sn.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F,x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
set.seed(1)
umap_res <- umap(as.matrix(sn.df[,-543]), n_components =  3, y = sn.df[,543]) 

res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
out2 = do.olda(as.matrix(sn.df[,-543]), ndim = 3, class)
senator.olda = data.frame(out2$Y,class)
stars(senator.olda[,-4], senator.olda$class, col.palette = dk.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
out2 = do.ulda(as.matrix(sn.df[,-543]), ndim = 3, class)
senator.ulda = data.frame(out2$Y,class)
stars(senator.ulda[,-4], senator.ulda$class, col.palette = dk.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">

```{r}
load("datasets/senatordf.rda")
res1 = viz2d_3d(sndf[, -ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[, ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = senator.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>

<td align="center" valign="center">
```{r}
load("datasets/senatordf.rda")
radialvis3d(data = sndf[, -ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = senator.palette, class.labels = c("Democratic", "Republican"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>



## Fig. S7: Autism Spectrum Disorder (ASD) Screening (Autisim Dataset)

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/Autism_normal.rda")
sn <- as.matrix(Autism_normal_class[,-1])
class <- Autism_normal_class$class
sn.df <- data.frame(sn,class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.df[,-15]), dims = 3, 
              perplexity=min(500,floor((nrow(sn.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>


<td align="center" valign="center">
```{r}
set.seed(1)
umap_res <- umap(as.matrix(sn.df[,-15]), n_components =  3, y = sn.df[,15]) 

res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
out2 = do.olda(as.matrix(sn.df[,-15]), ndim = 3, class)
autism.olda = data.frame(out2$Y,class)
stars(autism.olda[,-4], autism.olda$class, col.palette = dk.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
out2 = do.ulda(as.matrix(sn.df[,-15]), ndim = 3, class)
autism.ulda = data.frame(out2$Y,class)
stars(autism.ulda[,-4], autism.ulda$class, col.palette = dk.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
load("datasets/autismdf.rda")
res1 = viz2d_3d(sndf[, -ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[, ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = senator.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>

<td align="center" valign="center">
```{r}
load("datasets/autismdf.rda")
radialvis3d(data = sndf[, -ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = senator.palette, class.labels = c("Normal", "ASD subject"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>

## Fig. S8: SPECT Heart Datset

<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/SPECT_normal.rda")
sn <- as.matrix(SPECT_normal_class[,-1])
class <- as.factor(SPECT_normal_class$class)
sn.df <- data.frame(sn,class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.df[,-23]), dims = 3, 
              perplexity=min(500,floor((nrow(sn.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
set.seed(1)
umap_res <- umap(as.matrix(sn.df[,-23]), n_components = 3, y = sn.df[,23])
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = dk.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
out2 = do.olda(as.matrix(sn.df[,-23]), ndim = 3, class)
spect.olda = data.frame(out2$Y,class)
stars(spect.olda[,-4], spect.olda$class, col.palette = dk.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
out2 = do.ulda(as.matrix(sn.df[,-23]), ndim = 3, class)
spect.ulda = data.frame(out2$Y,class)
stars(spect.ulda[,-4], spect.ulda$class, col.palette = dk.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
load("datasets/spectdf.rda")
res1 = viz2d_3d(sndf[, -ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[, ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = senator.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>


<td align="center" valign="center">
```{r}
load("datasets/spectdf.rda")
radialvis3d(data = sndf[, -ncol(sndf)], cl = sndf[,ncol(sndf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = senator.palette, class.labels = c("Normal", "Abnormal"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>




## Fig. S8: Indic scripts (Script Dataset, 6 groups)
<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/indic_normal.rda")
sn <- as.matrix(indic_normal_class[,-1])
class <- as.factor(indic_normal_class$class)
script <- data.frame(sn,class)
sn.df = script
g = c(1,3,4,5,6,11)
sn.g.df = sn.df[sn.df$class %in% g, ]
class = droplevels(sn.g.df$class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.g.df[,-117]), dims = 3, 
              perplexity=min(500,floor((nrow(face.g.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")

# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = script.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
set.seed(1)
umap_res <- umap(as.matrix(sn.g.df[,-117]), n_components = 3, y = sn.g.df[,117])
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")

# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = script.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
out2 = do.olda(as.matrix(sn.g.df[,-117]), ndim = 4, class)
script.olda = data.frame(out2$Y,class)
stars(script.olda[,-5], script.olda$class, col.palette = script.palette)
```
<br />
(c) Stars with OLDA
</td>

<td align="center" valign="center">
```{r}
out2 = do.ulda(as.matrix(sn.g.df[,-117]), ndim = 4, class)
script.ulda = data.frame(out2$Y,class)
stars(script.ulda[,-5], script.ulda$class, col.palette = script.palette)
```
<br />
(d) Stars with ULDA
</td>
</tr>

<tr>
<td align="center" valign="center">
```{r}
load("datasets/scriptdf.rda")
res1 = viz2d_3d(scriptdf[, -ncol(scriptdf)])
res1 = data.frame(res1, as.factor(scriptdf[, ncol(scriptdf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = script.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(e) Viz3D
</td>



<td align="center" valign="center">
```{r}
load("datasets/scriptdf.rda")
radialvis3d(data = scriptdf[, -ncol(scriptdf)], cl = scriptdf[,ncol(scriptdf)], domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = script.palette, class.labels = c("Bangla", "Gujarati", "Gurmukhi", "Kannada", "Malayalam", "Urdu"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(f) RadViz3D
</td>
</tr>
</table>

## Fig. S9: RNA sequences of human tissues (Gtex Dataset)
<table width="500" border="0" cellpadding="0">

<tr>

<td align="center" valign="center">
```{r}
load("datasets-gdt/GTEX_normal.rda")
sn <- as.matrix(d.red_normal)
class <- as.factor(as.numeric(as.factor(group)))
class = as.factor(gsub("13","3",class))
sn.df = data.frame(sn,class)
g = c(2,4,5,7,8,9,11,12)
sn.g.df = sn.df[sn.df$class %in% g, ]
class = droplevels(sn.g.df$class)
set.seed(1)
tsne <- Rtsne(as.matrix(sn.g.df[,-20243]), dims = 3, 
              perplexity=min(500,floor((nrow(face.g.df)-1)/3)), 
              check_duplicates = F,  
              verbose=F, max_iter = 500)
res1 = tsne$Y
res1 = data.frame(res1, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = gtex.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(a) t-SNE
</td>

<td align="center" valign="center">
```{r}
umap_res <- umap(as.matrix(sn.g.df[,-20243]), n_components = 3, y = sn.g.df[,20243], target_weight = 0.5)
# umap_res <- umap(as.matrix(sn.g.df[,-20243]), n_components = 3)
res1 = data.frame(umap_res, class)
colnames(res1) = c("x","y","z","class")
# open 3d window
 
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75, 
                  col = gtex.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
rglwidget()
```
<br />
(b) UMAP
</td>
</tr>


<!-- ### Stars with OLDA -->
<!-- ```{r} -->
<!-- out2 = do.olda(as.matrix(sn.g.df[,-20243]), ndim = 4, class) -->
<!-- rna.olda = data.frame(out2$Y,class) -->
<!-- stars(rna.olda[,-5], rna.olda$class, col.palette = script.palette) -->
<!-- ``` -->

<!-- ### Stars with ULDA -->
<!-- ```{r} -->
<!-- out2 = do.ulda(as.matrix(sn.g.df[,-20243]), ndim = 4, class) -->
<!-- rna.ulda = data.frame(out2$Y,class) -->
<!-- stars(rna.ulda[,-5], rna.ulda$class, col.palette = script.palette) -->
<!-- ``` -->

<tr>
<td align="center" valign="center">

```{r}
load("datasets/gtexdf.rda")
res1 = viz2d_3d(sndf[, -ncol(sndf)])
res1 = data.frame(res1, as.factor(sndf[, ncol(sndf)]))
colnames(res1) = c("x","y","z","class")
# open 3d window
# resize window
par3d(windowRect = c(100, 100, 612, 612))
# plot
with(res1, plot3d(axes = F, x,y,z, type="s",size = 0.75,
                  col = gtex.palette[1:nlevels(class)][as.numeric(class)]))
box3d()
# add legend
# legend3d("topright", legend = paste('Group',1:nlevels(res1$class)),
#          pch = 16, col = df.palette[1:nlevels(res1$class)],
#          cex=1, inset=c(0.05))
rglwidget()

```
<br />
(c) Viz3D
</td>

<td align="center" valign="center">
```{r}
load("datasets/gtexdf.rda")
radialvis3d(data = sndf[, -ncol(sndf)], cl = as.factor(sndf[,ncol(sndf)]), domrp = F, doGtrans = F, lwd = 2, alpha = 0.35, pradius = 0.02, color = gtex.palette, class.labels = c("Breast", "Colon", "Esophagus", "Liver", "Lung", "Prostate", "Stomach", "Thyrioid"))
# rgl::par3d(windowRect = c(0, 0, 1200, 1200))
# rgl::par3d(viewport = c(0,0,1200, 1200))
# legend3d("topright", legend = c("Normal", "ASD subject"), pch = 19, col = senator.palette, cex=0.8, inset=c(0.01))
rgl::rgl.viewpoint(zoom = 0.6)
rgl::rglwidget()
```
<br />
(d) RadViz3D
</td>
</tr>
</table>










